---
title: "Story_6_pitch"
author: "Declan Oberlies"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
library(readr)
library(plotly)
```

# “Water is life”

For my final data story, I will be looking at SDG 6: Ensure availability and sustainable management of water and sanitation for all. 

First I will use this [data set](https://ourworldindata.org/grapher/proportion-using-safely-managed-drinking-water?time=2001..2024) to find what countries are most burdened by unsafe drinking water. I will then look at the disparities between urban and rural water from [this dataset](https://ourworldindata.org/grapher/urban-vs-rural-safely-managed-drinking-water-source?time=2024)
```{r}
# load data ----
url <- "https://ourworldindata.org/grapher/urban-vs-rural-safely-managed-drinking-water-source.csv?v=1&csvType=full&useColumnShortNames=true"
UR <- read.csv(url)

# modify data ----
UR <- UR %>% 
  mutate(RvU_deficite = wat_sm__residence_rural - wat_sm__residence_urban)

# plot data ----
UR <- UR %>% 
  filter(entity == "World")
ggplot(UR,
       aes(x = year,
           y = RvU_deficite)) +
  geom_line() +
  scale_y_reverse() +
  labs(title = "Fig.1")

```
From this preliminary graph(fig.1), we can see that the difference in water quality/availability between urban and rural populations has decreased significantly in recent years, and trends towards equality. This is one data story that I will explore more in depth for specific countries.

I will then use some of them many data sets from [here](https://ourworldindata.org/water-use-stress) to examine how freshwater is distributed within countries. 
```{r}
# load data ----
url <- "https://ourworldindata.org/grapher/freshwater-withdrawals-as-a-share-of-internal-resources.csv?v=1&csvType=full&useColumnShortNames=true"
withdrawSus <- read.csv(url)
# modify data ----
withdrawSus <-
  withdrawSus %>%
  rename(`sustainable withdraws` = X_6_4_2__er_h2o_stress__no_breakdown)

# plot data ----
ggplotly(
  ggplot(withdrawSus,
       aes(x = entity,
           y = `sustainable withdraws`,
           group = entity)) +
  geom_point(size = .5) +
    labs(title = "Fig.2")
  
  # +
  # geom_boxplot()
  )
```
This preliminary graph (Fig.2) shows the percentage of fresh water that countries can withdraw sustainably that they are withdrawing. Interestingly, countries like Kuwait are able to withdraw over 100% because they use desalination plants. This is a data story that I will explore more, and there are many more great data stories to be discovered from this dataset. 


I hypothesize that there will be situations where one country has limited water to clench the thrust of their people but uses lots of water for agriculture or industry. If this is the case then I will explore the dilemma. People need to eat and have jobs from industry but also need clean water. I think it will be interesting to see how different countries balance this.

# more stuff if I have time
I would like to look at ground water vs surface water data, but have not yet found a good dataset for that. I also want to look at deaths from unsafe/lack of water from [this dataset](https://ourworldindata.org/grapher/share-deaths-unsafe-water) but struggling to download the data. If I get it figured out then I will include it.


